Parsing Postfix log files

### 2009-02-26 doc/logparser-architecture.tex
Proofread the glossary and made corrections. Removed some unnecessary glossary entries for acronyms.
Checks for lines containing only spaces and "log line" in the architecture chapter.
Corrections from proofreading for first couple of sections.

To avoid cluttering the explanation of the parser architecture with the
details involved in parsing Postfix log files and the description of the
parser implemented for this project, the two have been separated.  This
chapter presents the architecture designed and developed for this project,
beginning with the overall architecture and design, followed by the three
components of the architecture: Framework, Actions, and Rules, and
finishes by describing the characteristics of this architecture.

@@

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It provides the parsing loop, shared data
        storage, loading and validation of rules, storage of results, and
        other support functions.

    \item [Actions] Each action performs the work required to deal with a
        single category of inputs, e.g.\ processing data from rejections.
        Actions are invoked to deal with a log line once it has been
        identified by the rules: actions modify data structures, handle
        complications, and cause data to be saved to the database.

    \item [Rules]  The rules are responsible for classifying inputs; they
        specify the action to invoke and the regex that matches the inputs
        and extracts data.  Rules provide an easily extensible method of
        associating log lines with actions.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework provides services to the
actions, but does not need to perform any tasks specific to the input being
parsed; actions benefit from having services provided by the framework,
freeing them to concentrate on the task of accurately and correctly
processing the information provided by rules; rules handle the low level
details of classifying inputs and extracting data from those inputs.

Separating the rules from the actions and framework makes it possible to
parse new log lines without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to match the log lines is
trivial in comparison to understanding a program's entire parser,
identifying the correct location to change, and making the appropriate
changes.  Bear in mind that changes to the parser must be made without
adversely affecting existing parsing, particularly as there may be edge
cases that are not immediately obvious ---
\sectionref{yet-more-aborted-delivery-attempts} describes a complication
that occurs only four times in \numberOFlogFILES{} log files.  Requiring
changes to the parser's source code also complicates upgrades, as the
changes must be preserved during the upgrade, and may clash with changes
made by the developer.  This architecture allows the user to add new rules
without changing the parser, unless the new log lines require functionality
not already provided by the existing actions.  If the new log lines do
require new functionality, new actions can be added to the parser without
modifying existing actions (\sectionref{adding new actions in
implementation} describes how to safely add new actions); only in the rare
event that the new actions require support from other sections of the code
will more extensive changes be required.

There is some similarity between the parser's design and William Wood's
\acronym{ATN}~\cite{atns}, used in Computational Linguistics for creating
grammars to parse or generate sentences.  The resemblance between the two
(shown in \tableref{Similarities between ATN and this architecture}) is
accidental, but it is obvious that the two different approaches share a
similar division of responsibilities, despite having different semantics.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Overall       & Determines the sequence 
                                          of transitions              \\
                        & Algorithm     & or actions that 
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assembles data and
                                          imposes conditions          \\
                        &               & the input must meet to be
                                          accepted as                 \\
                        &               & valid.                      \\
        Abbreviations   & Rules         & Responsible for 
                                          classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the programmers writing actions to concentrate
on writing productive code.  It links actions and rules, allowing either to
be improved independently of the other.  It provides shared storage to pass
data between actions, loads and saves state, loads and validates rules,
manages parsing, invokes actions, tracks how often each rule matches to
optimise rule ordering (\sectionref{rule ordering for efficiency}), stores
results of parsing, and miscellaneous other tasks.
Most parsers will require the same basic functionality from the framework,
plus some specialised support functions.  The framework is the core of the
architecture and is deliberately quite simple: the rules deal with the
variation in inputs, and the actions deal with the intricacies and
complications encountered when parsing.

The function that finds the rule matching the input and invokes the
requested action can be expressed in pseudo-code (indentation denoting flow
of control) as:

% DO NOT REFORMAT!

>>>>

To avoid cluttering the explanation of the parser architecture with the
details involved in parsing Postfix log files and the description of the
parser implemented for this project, the two topics have been separated.
This chapter presents the architecture designed and developed for this
project, beginning with the overall architecture and design, followed by
the three components of the architecture: Framework, Actions, and Rules,
and finishes by describing the characteristics of this architecture.

@@

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It manages the parsing process, providing shared
        data storage, loading and validation of rules, storage of results,
        and other support functions.

    \item [Actions] Each action performs the work required to deal with a
        single category of inputs, e.g.\ rejecting a delivery attempt.
        Actions are invoked to process an input once it has been classified
        by a rule.

    \item [Rules]  The rules are responsible for classifying inputs; they
        specify the action to invoke and the regex that matches the inputs
        and extracts data.  Rules provide an easily extensible method of
        associating inputs with actions.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework provides services to the
actions, but does not need to perform anything specific to the input being
parsed; actions benefit from having services provided by the framework,
freeing them to concentrate on the task of accurately and correctly
processing the inputs and the information provided by rules; rules handle
the low level details of classifying inputs and extracting data from those
inputs, for processing by actions.

Separating the rules from the actions and framework makes it possible to
parse new inputs without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to match the inputs is
trivial in comparison to understanding an entire parser, identifying the
correct location to change, and making the appropriate changes.  Bear in
mind that changes to a parser must be made without adversely affecting
existing parsing, particularly as there may be edge cases that are not
immediately obvious --- \sectionref{yet-more-aborted-delivery-attempts}
describes a complication that occurs only four times in \numberOFlogFILES{}
log files.  Requiring changes to a parser's source code also complicates
upgrades, as the changes must be preserved during the upgrade, and may
clash with changes made by the developer.  This architecture allows the
user to add new rules without editing a parser, unless the new inputs
require cannot be processed by the existing actions.  If the new inputs do
require new functionality, new actions can be added to the parser without
having to modify existing actions; only in the rare event that the new
actions require support from other actions will more extensive changes be
required.

There is some similarity between this architecture and William Wood's
\acronym{ATN}~\cite{atns,nlpip}, used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
the two (shown in \tableref{Similarities between ATN and this
architecture}) is accidental, but it is clear that the two different
approaches share a similar division of responsibilities, despite having
different semantics.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Parser        & Determines the sequence 
                                          of transitions              \\
                        &               & or actions that 
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assemble data and
                                          impose conditions           \\
                        &               & the input must meet to be
                                          accepted as                 \\
                        &               & valid.                      \\
        Abbreviations   & Rules         & Responsible for 
                                          classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

The framework takes care of miscellaneous support functions and managing
the parsing process, freeing the programmers writing actions to concentrate
on writing productive code.  It links actions and rules, allowing either to
be improved independently of the other.  It provides shared storage to pass
data between actions, saves and loads state, loads and validates rules,
manages parsing, invokes actions, tracks how often each rule matches to
optimise rule ordering (\sectionref{rule ordering for efficiency}), stores
results of parsing, and miscellaneous other tasks.  Most parsers will
require the same basic functionality from the framework, plus some
specialised support functions XXX EXPAND\@.  The framework is the core of
the architecture and is deliberately quite simple: the rules deal with the
variation in inputs, and the actions deal with the intricacies and
complications encountered when parsing.

The function that finds the rule matching the input and invokes the
requested action can be expressed in pseudo-code (with indentation denoting
flow of control) as:
